---
title: "Problem Set 4"
author: "Raj Thakkar"
date: "February 20, 2019"
output:
  html_document:
    df_print: paged
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, tidy = TRUE)
```

#### Note: I worked with Bhavna Sinha and Varun Miranda from the class

### Using loess or otherwise, fit a model to predict log10(budget) from year and length. For simplicity, do not transform year and length (even though a transformation of length would probably be sensible.)

### Should you fit a linear or curved function for year?
### Should you fit a linear or curved function for length?
### Do you need an interaction between year and length?
### What span should you use in your loess smoother?
### Should you fit using least squares or a robust fit?
### Some of these choices are clear-cut, while others will be a matter of preference. Either way, you must justify all your choices.

```{r}
library(tidyverse, quietly = TRUE)
cb_palette = c("#999999", "#E69F00", "#56B4E9", "#009E73", "#F0E442", "#0072B2",
"#D55E00", "#CC79A7")
movie_budgets <- read.csv("movie_budgets.txt", sep="")
movie_budgets$log_budget <- log10(movie_budgets$budget)

```

1. When we plot log_budget against year conditional on length, we observe that log_budget variable decreases with increase in year when length <= 72, it's almost constant with respect to year for lengths between 72 and 86 and then the slope becomes positive after length = 86 i.e. the variable log_budget increases with increase in year. The slope also increases as the values of length increase. Thus we need to fit a curved function for year conditional on length.

```{r}
# Code to plot log_budget against year conditional on length
# Code has been commented out and is shown only for reference

# gg = ggplot(movie_budgets, aes(x = year, y = log_budget)) + geom_point() +  geom_smooth(method = 'loess', method.args = list(degree = 1,family = "symmetric"),span = 1) +
 # facet_wrap(~cut_number(length, n = 8), ncol = 4)
# gg + labs(title = "Movie Budgets data split by length", subtitle = "Loess model with degree = 1, span = 1 and 'symmetric' family (in blue)") + xlab("Year") + ylab("Movie Budget (US dollars, log10 scale)")

```



2. We observe that log_budget increases with increase in length but the relationship is non monotonic as the log_budget variable becomes almost constant or it decreases after length = 200 conditional on year. Because of this we will ft a curved model (loess) for length conditional on year.  The non monotonic of relationship between log_budget and length is only because of the outliers which are movies with length > 200. The curves are similar for all the years except [1906,1955] and (1955,1975] which have outliers (movies with length > 200). The span of 1 gives us an ideal fit which is not affected a lot by the outliers (movies with length > 200)

```{r}
# Code to plot log_budget against length conditional on year
# Code has been commented out and is shown only for reference

# gg = ggplot(movie_budgets, aes(x = length, y = log_budget)) + geom_point() +  geom_smooth(method = 'loess', method.args = list(degree = 1,family = "symmetric"),span = 1) +
# facet_wrap(~cut_number(year, n = 8), ncol = 4)
# gg + labs(title = "Movie Budgets data split by year", subtitle = "Loess model with degree = 1, span = 1 and 'symmetric' family (in blue)") + xlab("Length of movie (in minutes)") + ylab("Movie Budget (US dollars, log10 scale)")

```


3. We observe that the shape of curve fitted for log_budget vs length with different cuts on years are almost similar except when there are outliers i.e. the movies with length > 200. But when we look at the shape of curve fitted for log_budget vs year with different cuts on lengths, the slope of the curve goes from negative to positive. Because of this relationship we need to have an interaction in our model.

4. We have used a span of 1 for loess as we wanted that the curve should not be affected a lot by the trend in outliers i.e movies with length > 200.

5. Since, we have outliers in the dataset, we have used "symmetric" family in loess which is a robust fit version of loess.

6. Since correlation between log_budget and year is very low (-0.0818) while the correlation is better between log_budget and length (0.634). We will use length to predict the log_budget conditional on year.

### 2. Draw ONE set of faceted plots to display the fit either condition on year or length, whichever seems to you to be more interesting. Choosing a sensible number of panels. Briefly describe what this set of plots shows you.



We want to plot the fit against length conditional on year. This requires a grid which is dense in length and sparse in year.

```{r}

budget.lo = loess(log_budget ~ length * year, data = movie_budgets, degree = 1, family = "symmetric",span = 1)

movies.grid = expand.grid(year = c(1906, 1942, 1975, 1996, 2005), length = seq(1,
390, 1))
movies.predict = predict(budget.lo, newdata = movies.grid)

movies.predict = data.frame(movies.grid, predicted_log10_budget = as.vector(movies.predict))

ggplot(movies.predict,aes(x = length, y = predicted_log10_budget, group = year, color = factor(year))) +
 geom_line() + scale_color_manual(values = cb_palette) + labs(color = "Year") + labs(title = "Movie Budget fit using length conditional on year", subtitle = "Loess model with interaction, degree = 1, span = 1 and family = 'symmetric'  ") + xlab("Length of movie (in minutes)") + ylab("Predicted Movie Budget (US dollars, log10 scale)")

```



Observations from the faceted plots of  the fit of Movie Budget in log10 scale conditional on year:

1. The relationship between movie budget in log10 scale and length is not monotonic. The rate of increase in budget (slope) decreases for movies with length > 200. The curves don't seem to converge stating that year of release always matters when we are predicting the movie budget on log10 scale using length.

### 3. Draw a raster-and-contour plot (or other 3D plot of your choice) to further display your fit. Briefly describe what, if anything, this plot shows you that your plot for question 2 didn't.

```{r}
budget.lo = loess(log_budget ~ length * year, data = movie_budgets, degree = 1, family = "symmetric",span = 1)

movies.grid = expand.grid(year = seq(1906, 2005, 1), length = seq(1,
390, 1))
movies.predict = predict(budget.lo, newdata = movies.grid)




# Dataframe
movies.plot.df = data.frame(movies.grid, fit = as.vector(movies.predict))

ggplot(movies.plot.df, aes(x = length, y = year, z = fit)) + geom_raster(aes(fill = fit)) +
coord_fixed() + scale_fill_distiller(palette = "RdYlBu") + geom_contour() + labs(title = "Raster and contour plot of fit of the model with year and length", subtitle = "Fit is for loess model with interaction, degree = 1, span = 1 and 'symmetric' family") + xlab("Length of movie (in minutes)") + ylab("Year")

```

Observations from the raster and contour plot of fit of Movie Budget in log10 scale:

1. We see that model captures the fact that the budget on log10 scale of a movie doesn't increase a lot when the length increases beyond 200.