---
title: "Problem Set 5"
author: "Raj Thakkar"
date: "February 26, 2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, tidy = TRUE)
```

### 1. Fit a model that estimates average BMI as a function of height and age. Note that a purely linear model is unlikely to be adequate. Write down the line of code you used to fit your model.

```{r message = FALSE, warning=FALSE}

library(NHANES)
library(tidyverse)

cb_palette = c("#999999", "#E69F00", "#56B4E9", "#009E73", "#F0E442", "#0072B2",
"#D55E00", "#CC79A7")

nhanes_under_17 <- NHANES %>% filter(Age <= 17 & !is.na(Height) & !is.na(BMI)) %>% select(ID,Age, Height, BMI)


```


```{r}
# Cuts on Age
gg = ggplot(nhanes_under_17, aes(x = Height, y = BMI, color = as.factor(Age))) + geom_point() +  geom_smooth(method = "loess", span = 1,method.args = list(degree = 1), se = FALSE) +
facet_wrap(~cut_number(Age, n = 3), ncol = 3)
gg + labs(title = "BMI vs Height fo Children under 17 split by year")

```

```{r}

gg = ggplot(nhanes_under_17, aes(x = Height, y = BMI)) +
geom_point() + geom_smooth(method = "loess", method.args = list(degree = 1),
se = FALSE) + geom_smooth(method = "loess", color = "orange", se = FALSE) +
geom_smooth(method = "gam", formula = y ~ s(x), color = "grey",
se = FALSE) + facet_wrap(~cut_number(Age, n = 3), ncol = 3)
gg + labs(subtitle = "Blue is degree 1 loess, orange is degree 2 loess, grey is GAM")


```

We will use degree = 1 and span = 1/3 for loess as the residuals have almost no trend when we use these parameters




```{r} 

bmi.lo <- loess(BMI ~ Height*Age ,data = nhanes_under_17, degree = 1, span = 1)


library(broom)
bmi.lo.df = augment(bmi.lo)
gg = ggplot(bmi.lo.df, aes(x = Height, y = .resid)) + geom_point()
gg + geom_smooth() + facet_wrap(~cut_number(Age, n = 3), ncol = 3)


```

```{r}

ggplot(bmi.lo.df, aes(x = .fitted, y = abs(.resid))) + geom_point() + geom_smooth(method = "gam",
formula = y ~ s(x))


```

```{r}

var(bmi.lo.df$.fitted)/var(nhanes_under_17$BMI)
var(bmi.lo.df$.fitted)/(var(bmi.lo.df$.fitted) + var(bmi.lo.df$.resid))

```





```{r}

# bmi.lo = loess(BMI ~ Height * Age, data = nhanes_under_17, span = 0.2, parametric = "Age",
# drop.square = "Age")
# 
# library(broom)
# bmi.lo.df = augment(bmi.lo)
# gg = ggplot(bmi.lo.df, aes(x = Height, y = .resid)) + geom_point()
# gg + geom_smooth(method = "gam", formula = y ~ s(x)) 


```

```{r}
var(bmi.lo.df$.fitted)/var(nhanes_under_17$BMI)
var(bmi.lo.df$.fitted)/(var(bmi.lo.df$.fitted) + var(bmi.lo.df$.resid))

```


### 2. Draw a graph faceted by age that shows how, according to your model, average BMI varies with height.

We want to plot the fit against Height conditional on Age This requires a grid which is dense in Height and sparse in Age

```{r}

bmi.grid = expand.grid(Age = seq(2,17,1), Height = seq(min(nhanes_under_17$Height),
max(nhanes_under_17$Height), 0.1))
bmi.predict = predict(bmi.lo, newdata = bmi.grid)

bmi.predict = data.frame(bmi.grid, predicted_BMI = as.vector(bmi.predict))

ggplot(bmi.predict,aes(x = Height, y = predicted_BMI)) + geom_line() + labs(title = "BMI fit using Height conditional on Age", subtitle = "Loess model with interaction, degree = 1, span = 1/3") + xlab("Height (in cm)") + ylab("Predicted BMI") + facet_wrap(~cut_number(Age, n = 3), ncol = 3)

```

### 3. Draw a graph which uses color to denote age that shows how, according to your model, average BMI varies with height. Your color-scheme must be legible and reasonably color-blind friendly.

```{r}

library(viridis)

bmi.grid = expand.grid(Age = seq(2,17,1), Height = seq(min(nhanes_under_17$Height),
max(nhanes_under_17$Height), 0.1))
bmi.predict = predict(bmi.lo, newdata = bmi.grid)

bmi.predict = data.frame(bmi.grid, predicted_BMI = as.vector(bmi.predict))

ggplot(bmi.predict,aes(x = Height, y = predicted_BMI,group = Age, color = Age)) + geom_line() + scale_color_viridis() + labs(title = "BMI fit using Height conditional on Age", subtitle = "Loess model with interaction, degree = 1, span = 1/3") + xlab("Height (in cm)") + ylab("Predicted BMI") 

```


### 4. Draw a graph that uses contours to show how average BMI varies with both height and age.

```{r}

ggplot(bmi.predict, aes(x = Height, y = Age, z = predicted_BMI)) + geom_raster(aes(fill = predicted_BMI)) +
coord_fixed() + scale_fill_distiller(palette = "RdYlBu") + geom_contour()

```